---
title: "SIPP Experimentation"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

# Setup 

This code taken from Census website, with some edits made to variables and file location.

```{r}
#Load the "data.table", "dplyr", and "bit64" libraries
require("data.table")
require("bit64")
require("dplyr")

#Read in the Primary Data file. Choose only the variables you want to read in in order to save on memory.
#This code assumes that your working directory is the same directory as the data.
ds <- c("/Users/mca141/Documents/Lumina/lumina-sipp/data/pu2020.csv")
pu <- fread(ds, sep = "|", select = c(
  
  #Common case identification variables
  'SSUID','PNUM','MONTHCODE','ERESIDENCEID','ERELRPE','SPANEL','SWAVE', 'TEHC_ST',
  
  #The base weight
  'WPFINWGT',
  
  #Common demographics variables, including age at time of interview (TAGE)
  #	and monthly age during the reference period (TAGE_EHC)
  'ESEX','TAGE','TAGE_EHC','ERACE','EORIGIN','EEDUC','ECERT', 'EPROCERT', 'EED_SCRNR',
  
  #Additional variables for analysis
  'TPTOTINC','EDEBT_ED','EJSEDDEBT','EOEDDEBT','TJSEDDEBTVAL','TOEDDEBTVAL','TDEBT_ED','TPEARN'))

#Make sure all the column names are upper-case
names(pu) <- toupper(names(pu))

```

Other libraries I might need

```{r}
library(tidyverse)
library(haven)
library(srvyr)
```

# Exploration of full data set

Mean total ed debt by education level

```{r}
pu |>
  filter(
    EED_SCRNR == 2, # Not currently enrolled
    !is.na(TDEBT_ED) # Total ed debt not NA
  ) |>
  group_by(EEDUC) |>
  summarize(mean(TDEBT_ED))

```

Age breakdown
```{r}
pu |>
  group_by(TAGE) |>
  summarize() |>
  print(n = 90)
```

TAGE is age by year, 0-89.

Month breakdown

```{r}
pu |>
  filter(!is.na(TDEBT_ED),
         !is.na(TOEDDEBTVAL),
         !is.na(TJSEDDEBTVAL)) |>
  group_by(MONTHCODE) |>
  summarize(sum(TDEBT_ED),
            "testsum" = sum(TOEDDEBTVAL) + sum(TJSEDDEBTVAL),
            sum(TOEDDEBTVAL),
            sum(TJSEDDEBTVAL)
            )
```

Reference month is recorded, but it doesn't seem to be used for education debt - all months have same number of observations and values.

Spouse stuff

```{r}
pu |>
  filter(!is.na(TDEBT_ED),
         !is.na(TOEDDEBTVAL)) |>
  group_by(MONTHCODE) |>
  summarize(sum(TDEBT_ED),
            "testdiff" = sum(TDEBT_ED) - sum(TOEDDEBTVAL),
            sum(TOEDDEBTVAL)
            )
```

Amounts owed jointly with spouse are smaller by a couple orders of magnitude (when summed).

Certs/educ

```{r}
pu |>
  # filter(EEDUC %in% c(40,41)) |>
  group_by(ECERT,EEDUC) |>
  summarize(n()) |>
  print(n=41)
```


# Create Dec-only data frame

```{r}
#| label: Create sippdec


sippdec <- pu |>
  filter(
    MONTHCODE == 12,
    !is.na(EEDUC)
  ) |>
  
  # Create 4 education groups
  mutate(
    educ_grp = case_when(
      EEDUC <= 39 ~ "HS or less",
      EEDUC %in% c(40,41) ~ "Some college",
      EEDUC == 42 ~ "AA",
      EEDUC == 43 ~ "BA",
      TRUE ~ "Graduate"
    ),
    
    # Add certificates
    educ_grp_cert = case_when(
      educ_grp %in% c("HS or less", "Some college") & ECERT == 1 ~ "Certificate",
      educ_grp %in% c("HS or less", "Some college") & EPROCERT == 1 ~ "Certificate",
      TRUE ~ educ_grp
    ),
    
    # Add age groups
    age_grp = case_when(
      TAGE_EHC >= 18 & TAGE_EHC <= 24 ~ "18 to 24",
      TAGE_EHC >= 25 & TAGE_EHC <= 34 ~ "25 to 34",
      TAGE_EHC >= 35 & TAGE_EHC <= 44 ~ "35 to 44",
      TAGE_EHC >= 45 & TAGE_EHC <= 64 ~ "45 to 64",
      TRUE ~ NA
    )
  )
  
```

Create 4 education groups (incorporated into "Create sippdec")
```{r}
#| eval: false
#| include: false


sippdec <- sippdec |>
  filter(!is.na(EEDUC)) |>
  mutate(
    educ_grp = case_when(
      EEDUC <= 39 ~ "HS or less",
      EEDUC %in% c(40,41) ~ "Some college",
      EEDUC == 42 ~ "AA",
      EEDUC == 43 ~ "BA",
      TRUE ~ "Graduate"
    )
  )
```

Add certificates (incorporated into "Create sippdec")
```{r}
#| eval: false
#| include: false

sippdec <- sippdec |>
  mutate(
    educ_grp_cert = case_when(
      educ_grp %in% c("HS or less", "Some college") & ECERT == 1 ~ "Certificate",
      educ_grp %in% c("HS or less", "Some college") & EPROCERT == 1 ~ "Certificate",
      TRUE ~ educ_grp
    )
  )

```

Create age groups (incorporated into "Create sippdec")
```{r}
#| eval: false
#| include: false


sippdec <- sippdec |>
  mutate(
    age_grp = case_when(
      TAGE_EHC >= 18 & TAGE_EHC <= 24 ~ "18 to 24",
      TAGE_EHC >= 25 & TAGE_EHC <= 34 ~ "25 to 34",
      TAGE_EHC >= 35 & TAGE_EHC <= 44 ~ "35 to 44",
      TAGE_EHC >= 45 & TAGE_EHC <= 64 ~ "45 to 64",
      TRUE ~ NA
    )
  )

```


## Convert to survey object

```{r}
sippdecsvy <- sippdec |> as_survey_design(weights = WPFINWGT)
```

# Debt by education level

Table: total loans owned in own name, disaggregated by education, not currently enrolled

```{r}

sippdecsvy |>
  filter(EED_SCRNR == 2,
         educ_grp_cert %in% c("AA", "BA", "Graduate", "Certificate"),
         TOEDDEBTVAL > 0,
         # TAGE_EHC > 24,
         # TAGE_EHC < 66
         ) |>
  group_by(educ_grp_cert) |>
  summarize(meddebt = survey_median(TOEDDEBTVAL))
  
```

# Debt by education level (detailed graduates)

```{r}
# Graduate degrees
sippdecsvy |>
  filter(EED_SCRNR == 2,
         educ_grp_cert == "Graduate",
         TOEDDEBTVAL > 0,
         ) |>
  group_by(EEDUC) |>
  summarize(meddebt = survey_median(TOEDDEBTVAL))

# Professional cert
sippdecsvy |>
  filter(EED_SCRNR == 2,
         EPROCERT == 1,
         TOEDDEBTVAL > 0,
         ) |>
  group_by(EPROCERT) |>
  summarize(meddebt = survey_median(TOEDDEBTVAL))
  
```

# Debt by age group

```{r}

sippdecsvy |>
  filter(EED_SCRNR == 2,
         educ_grp_cert %in% c("AA", "BA", "Graduate", "Certificate"),
         TOEDDEBTVAL > 0,
         !is.na(age_grp),
         ) |>
  group_by(age_grp) |>
  summarize(meddebt = survey_median(TOEDDEBTVAL))
  
```


# Dec 2019: descriptive statistics

Count of all observations
```{r}
sippdecsvy |> survey_count()
```

Mean and median age
```{r}
sippdecsvy |> summarize(survey_mean(TAGE_EHC))
# Mean = 46.2

sippdecsvy |> summarize(medage = survey_median(TAGE_EHC))
# Median = 46
```

Mean age (completers, not currently enrolled)
```{r}
sippdecsvy |>
  filter(
    EED_SCRNR == 2,
    educ_grp_cert %in% c("AA", "BA", "Graduate", "Certificate"),
         TOEDDEBTVAL > 0) |> summarize(meanage = survey_mean(TAGE_EHC))
# Mean = 39

sippdecsvy |>
  filter(
    EED_SCRNR == 2,
    educ_grp_cert %in% c("AA", "BA", "Graduate", "Certificate"),
         TOEDDEBTVAL > 0) |> summarize(medage = survey_median(TAGE_EHC))
# Median = 37
```


Gender breakdown
```{r}
sippdecsvy |> group_by(ESEX) |> summarize(survey_prop())
# 51.6% female, 48.4% male
```



